### 2020-04-17 intro.tex
public data?
updated dataclean
first changes reflecting meeting - more coming

The Cray XK7 Titan supercomputer was the \#1 system in the world for a
very long time, and has remained a critically important computer
system through the end of its life in the Summer of 2019. It defied
scale with 18,688 individual compute nodes and delivered tens of
billions of computing hours to the U.S. Oepartment of Energy
mission-critical programs for nearly 7 years.

@@

Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.
\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Daily volume of GPU swaps with labels on most frequent dates.}
  \label{fig:chronology}
\end{figure*}
After an initial fairly stable break-in period, the first rework cycle
begins in early 2013 and ends by May. \fix{Details?}. The second
rework cycle begins in late August of 2013 and ends by mid
December. This is followed by a very long and stable period until
about mid 2016, when GPU failures begin to rise. This results in the
final rework cycle, replacing 11,000 GPUs throughout all of 2017.

While the 2016 resistor failures were a surprise, there is a known
phenomenon \cite{referenceneeded} that describes a two phase process
that leads to failure. It also explains why there was a very stable
period preceeding the failures. There is specific decay associated
with the electronic components where, as these products age, they will
eventually fail thresholds for guard band margin and other things. In
a lot of cases, they will just fail as well. \fix{Give a reference and
  expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later.

>>>>

The Cray XK7 Titan supercomputer was the \#1 system in the world for a
very long time, and has remained a critically important computer
system through the end of its life in the Summer of 2019. It defied
scale with 18,688 individual compute nodes and delivered tens of
billions of computing hours to the U.S. Department of Energy
mission-critical programs for nearly 7 years.

@@

Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.
\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Volume of GPU swaps detected at inventories. The break-in
    period before 2014 involved working out mechanical assembly issues
    at PCIe connectors
    frequent dates.}
  \label{fig:chronology}
\end{figure*}
The first two rework cycles before 2014 can be characterized as a
break-in period on a new system that is first of its kind and pushes
many technological boundaries, in this case mainly regarding blade
mechanical assembly. The resulting field engineering generated high
GPU swap counts. Late 2013 begins a very long period of stable
operation with very few issues until about mid 2016, when GPU failures
begin to rise. This results in the final rework cycle, replacing
11,000 GPUs throughout much of 2017.

The failures were traced to a failing resistor on the circuitboard
(not the GPU iself) While the 2016 resistor failures were a surprise, there is a known
phenomenon \cite{referenceneeded} that describes a two phase process
that leads to failure. It also explains why there was a very stable
period preceeding the failures. There is specific decay associated
with the electronic components where, as these products age, they will
eventually fail thresholds for guard band margin and other things. In
a lot of cases, they will just fail as well. \fix{Give a reference and
  expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later. \fix{Will we (are we required) make the data publicly
  available?} The data used in this study, representing over 100,000
years of operation may be the largest publicly available GPU
reliability data set.
